An upper bound to gas storage and delivery via pressure-swing adsorption in porous materials

### 2020-07-10 gas-adsorption/rsc-gas-ads-complete.tex
made \V_opt paragraph have two separate ideas, instead of intertwining them which was confusing

A promising approach to densify natural
gas~\cite{makal2012methane,mason2014evaluating} and
hydrogen~\cite{suh2011hydrogen,garcia2018benchmark} for vehicular storage at
room temperature is through physical adsorption in nanoporous
materials~\cite{schoedel2016role}. The internal surfaces of porous materials
attract gas molecules through van der Waals, electrostatic, etc.\ interactions
to achieve a higher adsorbed gas density than the bulk gas at the same
temperature and pressure, allowing for room temperature and lower-pressure
storage and alleviating many drawbacks of high-pressure, low-temperature
storage.

@@

In this work, we place a rigorous upper bound on the deliverable capacity of a
pure gas in a rigid substrate by endowing a control volume with a spatially
uniform background energy field. Instead of using a molecular model for the
gas~\cite{gomez2017impact}, we use the experimental equation of state to
account for gas-gas interactions. In addition, we prove using the calculus of
variations that this spatially uniform substrate yields a maximal deliverable
capacity. We use our framework to place an upper bound on the deliverable
capacity of methane and hydrogen gas.

@@

Fig.~\ref{fig:methane-298-D} clearly shows that $\V_{opt}$ provides a maximum
$\rho_D$ over all \emph{spatially uniform}, static potential energy fields. To
qualitatively argue that the spatially uniform potential $\V_{opt}$ provides a
maximum deliverable capacity over \emph{all} (including non-spatially uniform)
static potential energy fields (a broader claim), consider a
non-spatially-uniform variation on $\V_{opt}$. The effect of
this variation is to reduce the deliverable capacity: At low adsorbed gas
densities, the lowest-energy positions will be preferentially occupied, while
at higher densities, the additional gas molecules will be forced into higher
energy locations. Thus, the mean attraction will be lowest at \pfull\ and
highest at \pempty. As a result, $\gst$ increases monotonically with
pressure. As shown in Sec.~\ref{sec:monotonic}, this results in
the deliverable capacity of a non-uniform potential with a given $\gst$ at $\pfull$ or $\pempty$ being lower than the
deliverable capacity of a uniform potential with $\V=\gst$. Thus, a spatially uniform potential gives the maximal
deliverable capacity amongst all static potential energy fields that give the
same $\gst$, and the \emph{optimal} uniform potential $\V_{opt}$ will yield a
greater deliverable capacity over any static, non-uniform potential.

@@

The grand canonical partition function of the control volume is:

\begin{multline}
    \Xi(\mu, V, T)= \\ \displaystyle \sum_{N=0}^\infty \frac{1}{\Lambda^{3N}N!} \int_{\Omega} \cdots \int_{\Omega} e^{-\beta E_{gg}(\rvec_1, ..., \rvec_N)} e^{\beta (\mu - \V) N} d\rvec_1 \cdots d\rvec_N.
    \label{eq:gcpf}
\end{multline}
We recognize this as the grand canonical partition function of the bulk gas in
the control volume without the external potential, but shifted in chemical
potential:

\begin{equation}
    \Xi(\mu, V, T)=\Xi_0(\mu - \V, V, T)
    \label{eq:xi_vs_xi0}
\end{equation}
where $\Xi_0(\mu, V, T)$ is the grand partition function of the bulk
gas in the control volume in the absence of an external potential. Importantly,
this equivalency depends on the potential $\V$ being spatially uniform.
Therefore, the thermodynamic properties of the gas atoms in the spatially
uniform external potential $\V$ (adsorbed in our idealized substrate)
are equivalent to the properties of the bulk gas at chemical potential $\mu-\V$
(where $T$ is held fixed). As $\V$ becomes more negative, corresponding to a
more attractive adsorbent, the thermodynamic properties of the adsorbed gas in
our ideal substrate are equivalent to the gas at a higher chemical potential.

>>>>

A promising approach to densify natural
gas~\cite{makal2012methane,mason2014evaluating} and
hydrogen~\cite{suh2011hydrogen,garcia2018benchmark} for vehicular storage at
room temperature is through physical adsorption in nanoporous
materials~\cite{schoedel2016role}. The internal surfaces of porous materials
attract gas molecules through van der Waals, electrostatic, etc.\ interactions
to achieve a higher adsorbed gas density than the bulk gas at the same
temperature and pressure, allowing for room temperature and lower-pressure
storage and alleviating many drawbacks of high-pressure or low-temperature
storage.

@@

In this work, we place a rigorous upper bound on the isothermal deliverable capacity of a
pure gas in a rigid substrate by endowing a control volume with a spatially
uniform background energy field. Instead of using a molecular model for the
gas~\cite{gomez2017impact}, we use the experimental equation of state to
account for gas-gas interactions. In addition, we prove using the calculus of
variations that this spatially uniform substrate yields a maximal deliverable
capacity. We use our framework to place an upper bound on the deliverable
capacity of methane and hydrogen gas.

@@

Fig.~\ref{fig:methane-298-D} clearly shows that $\V_{opt}$ provides a maximum
$\rho_D$ over all \emph{spatially uniform}, static potential energy fields. To
qualitatively argue that the spatially uniform potential $\V_{opt}$ provides a
maximum deliverable capacity over \emph{all} (including non-spatially uniform)
static potential energy fields (a broader claim), consider a
non-spatially-uniform variation on $\V_{opt}$. 
At low adsorbed gas densities, the lowest-energy positions will be preferentially occupied, while
at higher densities, the additional gas molecules will be forced into higher
energy locations. Thus, the mean attraction will be lowest at \pfull\ and
highest at \pempty. 
Therefore, the effect of a non-spatially uniform variation on $\V_{opt}$ is to reduce the deliverable capacity.
By the same argument, $\gst$ increases monotonically with
pressure. As shown in Sec.~\ref{sec:monotonic}, this results in
the deliverable capacity provided by a non-uniform potential with a given $\gst$ at $\pfull$ or $\pempty$ being lower than the
deliverable capacity provided by a uniform potential with $\V=\gst$. Thus, a spatially uniform potential gives the maximal
deliverable capacity amongst all static potential energy fields that give the
same $\gst$, and the \emph{optimal} uniform potential $\V_{opt}$ will yield a
greater deliverable capacity over any static, non-uniform potential.

@@

The grand canonical partition function of the control volume is:
\begin{multline}
    \Xi(\mu, V, T)= \\ \displaystyle \sum_{N=0}^\infty \frac{1}{\Lambda^{3N}N!} \int_{\Omega} \cdots \int_{\Omega} e^{-\beta E_{gg}(\rvec_1, ..., \rvec_N)} e^{\beta (\mu - \V) N} d\rvec_1 \cdots d\rvec_N.
    \label{eq:gcpf}
\end{multline}
We recognize this as the grand canonical partition function of the bulk gas in
the control volume without the external potential, but shifted in chemical
potential:
\begin{equation}
    \Xi(\mu, V, T)=\Xi_0(\mu - \V, V, T)
    \label{eq:xi_vs_xi0}
\end{equation}
where $\Xi_0(\mu, V, T)$ is the grand partition function of the bulk
gas in the control volume in the absence of an external potential. Importantly,
this equivalency depends on the potential $\V$ being spatially uniform.
Therefore, the thermodynamic properties of the gas atoms in the spatially
uniform external potential $\V$ (adsorbed in our idealized substrate)
are equivalent to the properties of the bulk gas at chemical potential $\mu-\V$
(where $T$ is held fixed). As $\V$ becomes more negative, corresponding to a
more attractive adsorbent, the thermodynamic properties of the adsorbed gas in
our ideal substrate are equivalent to the gas at a higher chemical potential.
